Reasoning Graphically about Quantum Computation

### 2008-03-07 papers/reasoning-graphically/paper.tex
added conclusion.
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1407 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

There are several foundational approaches to graph transformation,
including algebraic appoaches~\cite{corradini97algebraic}, node-label
controlled~\cite{Graphgrammars83}, matrix
based~\cite{DBLP:conf/gg/VelascoL06}, and programmed graph
replacement~\cite{progress97}. These provide general ways of
interpretating graph transformations which can then be implemented to
provide machinery for a specific application.  However, systems based
on these theories do not provide machinery for the semantics of
compact closed categories. Thus their notion of matching and
replacement do not guarentee well-typed results. Furthermore, they do
not provide machinery to for rewriting of graphs with elipses
notation.

The distinctive feature of our form of graph rewriting is that the
graphs capture the structural properties of compact closed categories.
In this respect, our graphs caputure the structural properties of
compact closed categories and rewriting is compositional: it preserves
the type of the rewritten subgraph. However, the various components of
our system can be seen as instantiations for in a general graph
rewriting system: our matching provides the embedding information and
the node-data matching of the object formalism provides application
conditions and the attribute transfer function.

@@

\section{Conclusions}
\label{sec:conclusions}

Ideas for further work
\begin{itemize}
\item simplification ordering
\item confluence arguments for graphs
\item formalise algorithms in a thm prover
\item richer graph structures - quantification over edges
\item richer expression language in vertices
\item completeness of rewrite rules with respect to wpHilb
\end{itemize}

>>>>

There are several foundational approaches to graph transformation,
including algebraic approaches~\cite{corradini97algebraic}, node-label
controlled~\cite{Graphgrammars83}, matrix
based~\cite{DBLP:conf/gg/VelascoL06}, and programmed graph
replacement~\cite{progress97}. These provide general ways of
interpretation graph transformations which can then be implemented to
provide machinery for a specific application.  However, systems based
on these theories do not provide machinery for the semantics of
compact closed categories. Thus their notion of matching and
replacement do not guarantee well-typed results. Furthermore, they do
not provide machinery to for rewriting of graphs with ellipses
notation.

The distinctive feature of our form of graph rewriting is that the
graphs capture the structural properties of compact closed categories.
In this respect, our graphs capture the structural properties of
compact closed categories and rewriting is compositional: it preserves
the type of the rewritten subgraph. However, the various components of
our system can be seen as instantiations for in a general graph
rewriting system: our matching provides the embedding information and
the node-data matching of the object formalism provides application
conditions and the attribute transfer function.

@@

\section{Conclusions and Further Work}
\label{sec:conclusions}

We have introduced a representation for graphs which can formally
characterise the ellipses notation used informally to represent certain
infinite families of graph rewrites, such as the Spider Theorem. This
representation provides the foundation for a simple meta-logic for
reasoning about models of compact closed categories. We illustrated
this by providing an account of quantum computation and showing how
computation can be performed. Having developed the basic
representational machinery and shown matching to be decidable, we are
left with several exciting avenues for further research. The most
immediate is to provide a full implementation - a partial one is
already
available~\footnote{\url{http://dream.inf.ed.ac.uk/projects/quantomatic}}.
Other areas of further work include considering confluence results for
sets of rewrite rules, even richer representations for graph-patterns
and completeness of the set of the rewrite rules for quantum
computation.

% simplification ordering
% \item confluence arguments for graphs
% \item formalise algorithms in a thm prover
% \item richer graph structures - quantification over edges
% \item richer expression language in vertices
% \item completeness of rewrite rules with respect to wpHilb
% \end{itemize}
